A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway using an accessible web application

### 2018-03-08 manuscript/manuscript.Rmd
initial edits cont. "TPL/TPR"

###Identification of new Loss-of-Function/Gain-of-Function mutations
To test the extent of redundancy/specialization between sister pairs, the Natural Variation Webtool allows for identifications of accessions that can be potentially used to investigate the phenotypic effects in loss of function variants. For example, it has been shown that IAA10 has a missense mutation in the EAR domain, which converts the second leucine in the LxLxL motif to a serine (L49S), a mutation likely to reduce the ability of the IAA to recruit TOPLESS co-repressor [@lee_biological_2016]. This mutation is found in three accessions in the southern UK. A second potentially stronger mutation, represented in one accession from northeastern Spain is a missense resulting in an early stop codon in the PB1 domain (need Supp. Fig?). Such a mutation is likely to prevent IAA10 from interacting with an ARF, rendering it impotent. These alleles would be strong candidates for loss-of-function mutants, allowing researchers to discover the function of IAA10. In particular, it would make possible a straightforward path for creating an iaa10/11 double mutant using CRISPR/Cas9 directed mutagenesis, allowing for examination of the extent of redundancy between these family members.

Several examples of possible gain-of-function degron missense mutations were found in IAA2, IAA3, IAA6, IAA7, IAA8, IAA13, IAA15, IAA18, IAA19, IAA27, IAA28, IAA29. Most degron mutations were found in one or two accessions represented in the database (Supplemental Table or Map?). Interestingly, almost all of the IAAs found to have accessions with degron missense mutations have a closely related sister, which may indicate a redundant function within the sister-pairs. IAA6 natural variation has two potentially dominant missense mutations found within the degron (Supp Fig. alignment), one previously believed to lead to the shy1-1 phenotype [@kim_two_1996]. However, this variant is widely represented in 31 accessions across europe (Supplemental Map Fig). This finding supports the proposition that iaa6[C78R] may not be the mutation causing the shy1-1 phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)). These accessions could be used to confirm whether the shy1-1 phenotype exists within these plants. Additionally, an accession was identified with a V77M mutation in the IAA6 degron that may be useful to explore potential iaa6 gain-of-function phenotypes.

###Identifying novel functional sequences in IAAs
An exciting possible use for the Natural Variation Webtool is to search for conserved, low-diversity sites that may represent possible sites of post-translational modification or protein-protein interaction. For example, recent detailed biochemical analysis identified putative lysine ubiquitination sites in members of the IAA6/19 sister pairs [@winkler_variation_2017]; Supp Fig IAA6/19 alignment). These sites show no natural sequence variation in Arabidopsis, while several other lysines are represented in the natural sequence diversity (Supp Figs alignment and map). In IAA6, there are 17 lysines of which 6 are putative ubiquitination sites, one is in the KR motif, while two others show natural variation. In IAA19, there are 15 total lysines, with 8 as putative ubiquitination sites, 2 in the KR motif and degron, and 1 that exhibits natural variation. Natural variation sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.

The members of the IAA8/9/27 group have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006]. IAA8/9 have  redundant roles in leaf serration [@koenig_auxin_2009], while IAA8 important in lateral roots [@arase_iaa8_2012]. IAA8 has been reported to uniquely localize in the cytosol [@arase_iaa8_2012] where it may interact with LSD1 protein [@coll_programmed_2011]. The N-terminal extension has a conserved region shared between IAA8 and IAA9 with low natural sequence diversity (Supp Fig IAA8/9 alignment) that may represent a site of novel functionality, possibly mediating interaction with LSD1. There are a few accessions with missense alleles in this region in both IAA8 and IAA9 that may be valuable in probing the functionality of this region (Supplemental Fig IAA8/9 map).

Members of the IAA17/7/14/16 group have a second EAR motif (LxLxL) with no observed missense mutations in IAA7/16/17 (Supp Fig IAA7/14/16/17 alignment). This observation supports the experimental evidence for IAA7 that the second EAR motif is functional: it can interact with TPR1 corepressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016]. For IAA14, we did observe two missense mutations that may be of interest to probe functionality of the IAA14 EAR motif: one in the central “L” (L41F) that is represented in only one accession, and the other in the second “x” (K40N) that is found in 12 accessions around Copenhagen (Supp Fig. IAA14 map).

>>>>

###Identification of new Loss-of-Function/Gain-of-Function mutations
Accessions with functional variants can be potentially used to investigate the extent of redundancy/specialization within sister pairs. For example, it has been shown that IAA10 has a missense mutation in the EAR domain, which converts the second leucine in the LxLxL motif to a serine (L49S), a mutation likely to reduce the ability of the IAA to recruit the TOPLESS co-repressor [@lee_biological_2016]. This mutation was found in three accessions from the southern UK. A second potentially stronger mutation, represented in one accession from northeastern Spain is a ~~missense~~ *frameshift?* resulting in an early stop codon in the PB1 domain (need Supp. Fig?). This likely prevents IAA10 from interacting with an ARF. ~~These alleles would be strong candidates for loss-of-function mutants, allowing researchers to discover the function of IAA10. In particular, it would make possible a straightforward path for creating an iaa10/11 double mutant using CRISPR/Cas9 directed mutagenesis, allowing for examination of the extent of redundancy between these family members.~~ *I don't think we can say this. There are SALK lines in the 5' UTR and first intron of IAA10 and IAA11.*

Several examples of possible gain-of-function degron missense variants were found in IAA2, IAA3, IAA6, IAA7, IAA8, IAA13, IAA15, IAA18, IAA19, IAA27, IAA28, IAA29. Most degron variants were only found in one or two accessions (Supplemental Table or Map?). ~Interestingly, almost all of the IAAs found to have accessions with degron missense mutations have a closely related sister, which may indicate a redundant function within the sister-pairs.~ *But we know that iaa6 iaa19 iaa5 triple doesn't have a phenotype.* IAA6 has two potentially dominant missense mutations found within the degron (Supp Fig. alignment), one previously believed to lead to the shy1-1 phenotype [@kim_two_1996]. Interestingly, this variant is widely represented in 31 accessions across europe (Supplemental Map Fig). This finding supports the proposition that iaa6[C78R] may not be the mutation causing the shy1-1 phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)). These accessions could be used to confirm whether iaa6[C78R] is responsible for the SHY1-1D phenotype. Additionally, an accession was identified with the IAA6-V77M variant, also in the degron, that may be useful to further explore iaa6 gain-of-function phenotypes.

###Identifying novel functional sequences in IAAs
An exciting possible use for the Natural Variation Webtool is to search for conserved, low-diversity sites that may represent possible sites of post-translational modification or protein-protein interaction. For example, recent detailed biochemical analysis identified putative lysine ubiquitination sites in members of the IAA6/19 sister pairs [@winkler_variation_2017]; Supp Fig IAA6/19 alignment). These sites are conserved across natural sequence variation, while several other lysine residues vary (Supp Figs alignment and map). In IAA6, there are 17 lysines of which 6 are putative ubiquitination sites, one is in the KR motif, while two others show natural variation. In IAA19, there are 15 total lysines, with 8 as putative ubiquitination sites, 2 in the KR motif and degron, and 1 that exhibits natural variation. **This sentence is not clear. Only 1 of the ubiquitination sites varies?** Natural sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.

The members of the IAA8/9/27 clade have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006]. IAA8/9 play redundant roles in leaf serration [@koenig_auxin_2009], while IAA8 important in lateral roots [@arase_iaa8_2012]. IAA8 has been reported to uniquely localize in the cytosol [@arase_iaa8_2012] where it may interact with LSD1 protein [@coll_programmed_2011]. The N-terminal extension has a conserved region shared between IAA8 and IAA9 with low natural sequence diversity (Supp Fig IAA8/9 alignment) that may represent a site of novel functionality, possibly mediating interaction with LSD1. There are a few accessions with missense alleles in this region in both IAA8 and IAA9 that may be valuable in probing the functionality of this region (Supplemental Fig IAA8/9 map).

Members of the IAA17/7/14/16 group have a second EAR motif (LxLxL) with no observed missense mutations in IAA7, 16, or 17 (Supp Fig IAA7/14/16/17 alignment). This observation supports the experimental evidence for IAA7 that the second EAR motif is functional: it can interact with TPR1 corepressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016]. For IAA14, we did observe two missense mutations that may be of interest to probe functionality of the IAA14 EAR motif: one in the central lysine (L41F) that is represented in only one accession, and the other in the second “x” (K40N) that is found in 12 accessions around Copenhagen (Supp Fig. IAA14 map).
